Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2025-01-08 doc/charting/Bush-2022.qmd
update data extraction from my studies

@@

-   Github - open-source code, version control, collaboration -
    analysing
-   Python - open-source programming - analysing
-   R - open-source programming - analysing
-   arXiv/bioRxiv/psyRxiv - preprint sharing - disseminating
-   OpenNeuro - data sharing - disseminating
-   Open Science Framework - preregistration - planning
-   clinicaltrials.gov - preregistration - planning

@@

### Governance

@@

@@

>>>>

## Characteristics of the paper

-   Type of paper (e.g., tips, example): example
-   Themes (e.g., tools, organization): tools
-   Other keywords (e.g., newcomers): workflow

@@

-   Github - open-source code, version control, collaboration -
    analysing
-   Python - open-source programming - analysing
-   R - open-source programming - analysing
-   arXiv/bioRxiv/psyRxiv - preprint sharing - disseminating
-   OpenNeuro - data sharing - disseminating
-   Open Science Framework - preregistration - planning
-   clinicaltrials.gov - preregistration - planning
-   FAIR (findability”, “accessibility”, “interoperability”, and “reusability”) -principles to organize data, not a tool per se, but I think they 
    need to be emphasized - data collection, organization and sharing - 
-   Brain imaging data structure (BIDS) - a common and standardized framework all scientists can work on - data collection and sharing
-   Containarized pipelines - code that allows easy reproduction - data analysis
-   ReproBIDS - dictionary with standardized terms for BIDS neuro data - planning, data collection, data deposition

@@

@@

-   The steps below were identified after: 1) assessing the crisis of reproducibility, 2) literature search (systematic review) #worth noting 
    since it pops up in so many papers as a previous step to implementation.

@@

### Centre workflow of practices to become more open (the ones above are more on the individual level)

-   publish code, publishing pre-prints, standardizing data to BIDS, transition to containarized pipelines, publish data in repositories, 
    establish data dictionaries, pre-registration,

### 2025-01-08 doc/paper.qmd

@@

We define open collaboration using the definition as found in Forte and
Lampe (2013):

@@

-   Aim

@@

The full protocol for this scoping review was uploaded to the Open Science Framework [@Johnston2022].

This is a scoping review, so we are following the framework described by
Arksey and O'Malley
([DOI:10.1080/1364557032000119616](https://doi.org/10.1080/1364557032000119616))
and by Levac, Colquhoun, and O'Brien
([DOI:10.1186/1748-5908-5-69](https://doi.org/10.1186/1748-5908-5-69))
as well as the guidelines outlined in the PRISMA-ScR
([DOI:10.7326/M18-0850](https://doi.org/10.7326/M18-0850) statement.

### Deviations from protocol
As with any research project, things evolve compared to what was
originally intended or planned. In this section we describe what was
changed from the original, uploaded protocol.

#### Document types

We originally intended on searching websites, blogs, and other online
resources, but it was very difficult to programmatically and
systemtically achieve this. So we ended up only searching for scholarly
work that was indexed in databases.

#### Source databases

@@

Given these challenges however, we unintentionally found
[OpenAlex](https://openalex.org/), which is a public database of
scholarly output that aggregates dozens of other source databases into
one, easy interface. This resource also has an R package called
`{openalexR}` [@Aria2024] with a very well designed interface, which
simplified or made redundant much of our code we wrote. It also made it
unnecessary to use the individual R packages for each source database
that we originally listed in the protocol. By using this archive, it
resolved all our challenges and barriers we had before. For a list of
where they get their source data from, see their ["About the
data"](https://help.openalex.org/hc/en-us/articles/24397285563671-About-the-data)
page.

#### Inclusion and exlucion criteria

@@

@@

### Formal sources

We developed the initial search strategy in consultation with a research
librarian. We will collect the data via systematic searches of
databases, as well as some hand searches.

We used OpenAlex is an open database with scholarly works from MAG,
Crossref, ORCID, ROR, DOAJ, Unpaywall, Pubmed, Pubmed Central, The ISSN
International Centre, Internet Archive, Web crawls, Subject-area and
institutional repositories from arXiv and Zenodo.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(stringr)
library(here)
source(here("R/search-terms.R"))
search_terms_inline <- str_replace_all(search_terms$general, "\\*", "\\\\\\*")
```

We used the following search terms:

```         
> `r search_terms_inline`
```

Inclusion criteria included any document where open collaboration
practices are not the primary focus. We will be guided by the definition
of open collaboration from Forte and Lampe (2013) mentioned above in
determining whether the records are relevant. Any published document
with reporting on current open collaboration practices. Any published
document with advice, guidance, tools, and/or recommendations for
improving open collaboration. Article language in English.

Exclusion criteria were documents that do not report on specific open
collaboration practices. Documents published \>5 years in the past
(defined as before xxxx) to capture current/most recent practices.

All data obtained from all searches in online databases will be exported
into text files or CVS files for further processing including removal of
duplicates.

At least two of us will extract data using a standardized and tested
template. Data regarding the data source (e.g., author, title,
publication year), open collaboration practices, and any other relevant
information, will be extracted. Extracted data will be summarized with
the descriptive analytical method described by Arksey and O'Malley
(2005). The descriptive analytical method is aimed at identifying and
summarizing different open collaboration practices.

### Informal sources

>>>>

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE)
library(stringr)
library(here)
source(here("R/openalex-search.R"))
# TODO: This doesn't work, Luke is fixing it in connection with making Quarto website
# search_terms_inline <- str_replace_all(search_terms$general, "\\*", "\\\\\\*")
```

@@

We define open collaboration using the definition as found in
[@Forte2013]:

@@

### Aim

@@

The full protocol for this scoping review was uploaded to the Open
Science Framework [@Johnston2022]. As with any research project, things
evolve compared to what was originally intended or planned. In this
section we briefly re-state what we described in the protocol and
especially we describe what was changed from the protocol.

This is a scoping review, so we followed the framework described in
[@Arksey2005, @Levac2010] as well as the guidelines outlined in the
PRISMA-ScR statement [@Tricco2018].

### Deviations and challenges

@@

We also originally intended on searching websites, blogs, and other
online resources, but it was very difficult to programmatically and
systemtically achieve this. So we ended up only searching for scholarly
work that were indexed in databases.

Given these challenges however, we unintentionally found
[OpenAlex](https://openalex.org/), which is a public database of
scholarly output that aggregates dozens of other source databases into
one, easy interface. This resource also has an R package called
`{openalexR}` [@Aria2024] with a very well designed interface, which
simplified or made redundant much of our code we wrote. It also made it
unnecessary to use the individual R packages for each source database
that we originally listed in the protocol. By using this archive, it
resolved all our challenges and barriers we had before. For a full list
of where they get their source data from, see their ["About the
data"](https://help.openalex.org/hc/en-us/articles/24397285563671-About-the-data)
page.

### Document selection

We developed the initial search strategy in consultation with a research
librarian. We collected the data via systematic searches of databases.
All authors were involved in reviewing the collected sources, from
reviewing the titles, to abstracts, and finally to full-text documents.

#### Information sources

We used OpenAlex, after encountering the challenges described above,
which is an open database with scholarly works from Microsoft Academic
Graph, Crossref, ORCID, ROR, DOAJ, Unpaywall, Pubmed, Pubmed Central,
the ISSN International Centre, Internet Archive, Web crawls,
subject-area and institutional repositories from arXiv and Zenodo.

#### Search terms

We used the following search terms when searching for scholarly work:

> `r search_terms_inline`

#### Inclusion and exclusion criteria

Inclusion criteria included any document where open collaboration
practices are not the primary focus. We relied on the definition of open
collaboration from [@Forte2013] in determining whether the records were
relevant. Any published document with reporting on current open
collaboration practices. Any published document with advice, guidance,
tools, and/or recommendations for improving open collaboration. Article
language in English. Exclusion criteria were documents that do not
report on specific open collaboration practices.

@@

#### Charting procedure

At least two of us extracted data using a standardized and tested
template. Data regarding the data source (e.g., author, title,
publication year), open collaboration practices, and any other relevant
information, will be extracted. Extracted data will be summarized with
the descriptive analytical method described in [@Arksey2005], which is
aimed at identifying and summarizing different open collaboration
practices.

@@

The start date remained 5 years from our end date, which would have been
approximately 2019-02-28.
